Determining sequence-dependent DNA oligonucleotide hybridization and dehybridization mechanisms using coarse-grained molecular simulation, Markov state models, and infrared spectroscopy

### 2021-07-27 cover_letter/response_letter_MJ_210721.tex

\textbf{Author Reply:}   We thank the reviewer for their close reading of the manuscript and are pleased that they find the work impactful and suitable for publication in JACS.

\textcolor{blue}{Major point.
1)      Many of the core conclusions about the hybridization and dehybridzation mechanisms are based on the designation of the relevant metastable states as H, 5S2, 3S2, 5S4, F4, and D. While the authors do an excellent job of demonstrating that 2-6 state models provide an accurate description of the dynamics, in the current form, the paper does not provide sufficient evidence to support their chosen representative microstates and associated nomenclature (the only evidence being the ``representative'' structures shown in Figures 2A, C). A given macrostate represents an ensemble of microstates and it should be demonstrated that a significant fraction of the microstates that have a high probability of belonging to a particular macrostate share the same or similar base-pairing pattern. This would better justify how the authors chose to describe these states as 5S2, 3S2, etc., and provide the reader with more confidence that the mechanisms follow from the simulations. At the least, the authors should describe how the representative microstates (depicted in Figure 2A, C) were obtained.
}

\textbf{Author Reply:} The designation and structural representation of the macrostates is crucial to the core conclusions of the manuscript and we concur that we can and should better expose for the reader how the representative microstates were selected, how representative they are of the microstate ensemble within the macrostate, and our analysis of the base-pairing patterns within each macrostate that motivated our naming scheme.

On p.~19, we have added the following text:

``...each of these macrostates. \rood{The representative microstates emblematic of each macrostate were selected randomly from those possessing high ($>$99\%) macrostate membership probabilities. Since the microstate ensemble exhibits a range of conformations, however, it is useful to characterize the degree of structural heterogeneity to determine the degree to which these microstates are emblematic of the distribution. In Fig.~S{\blauw{X}} we present a projection of our macrostates into two physical order parameters -- the degree of 3$^\prime$ shift and degree of 5$^\prime$ shift -- to provide an interpretable embedding of the macrostates that exposes base pairing patterns and structural heterogeneity within the microstate ensemble. In all cases, we find these microstate distributions to be relatively narrowly focused within the low-free energy core of each macrostate such that a single archetypal microstate is indeed a good representative for the ensemble and an accurate representation of the base pairing pattern of the macrostate.} In Fig. 2b we present...''

We also added the new Fig.~S7 \noteg{is this order fine?} to the Supporting Information where we discuss at some length within the caption the important issues raised by the reviewer. The figure and its caption are reproduced below.

\begin{figure}[ht!]
	\centering
    \includegraphics[angle=90,origin=c,width=1.0\textwidth]{FigS7.pdf}
\end{figure}

@@

\begin{figure}[ht!]
	\centering
    \includegraphics[angle=90,origin=c,width=1.0\textwidth]{FigS7.pdf}
\end{figure}

\begin{figure}[ht!]
	\begin{center}
        \caption*{\textbf{Figure S7.}\rood{Macrostate interpretation and structural analysis. The Markov state model macrostates are learned from the simulation trajectories using the featurization and deep learning procedure described in Section 2.1.2 of the main text, but it is informative to perform a \textit{post hoc} projection of the simulation data into a low-dimensional space spanned by physically interpretable order parameters in order to determine the structural characteristics of each macrostate and the structural heterogeneity of the microstate ensemble. (a) Definition of two physically-interpretable order parameters useful in characterizing the structure and base pairing of each microstate under both shifting and fraying. We define the $3^\prime$ shift for a particular base on a strand as the linear distance between that base and the base on the complementary strand offset by two base pairs in the $3^\prime$ direction. This distance reaches a minimum when the two strands are mutually shifted two base pairs in the $3^\prime$ direction. We define $d_{3^\prime}$ as the mean $3^\prime$ shift averaged over the two central bases, represented by red arrows, in the complementary strands. It is more convenient to work the reciprocal $d_{3^\prime}^{-1}$, since it has a finite range whereas $d_{3^\prime}$ diverges for dehybridized strands. We define $d_{5^\prime}^{-1}$ analogously as the reciprocal mean $5^\prime$ shift. We have found this pair of variables extremely informative in revealing the microstate and macrostate structure and base pairing patterns. (b) We report the mean and standard deviation of $d_{3^\prime}$ and $d_{5^\prime}$ taken over all microstates comprising each macrostate for each sequence. The small standard deviations that are consistent with a narrow distribution of the bulk of the microstates around the free energy minimum of each macrostate. (c) Superposition of sample density for each macrostate for each sequence AT-all, GC-end, GC-core, and GC-mix projected into $d_{3^\prime}^{-1}$-$d_{5^\prime}^{-1}$. This panel conveys two important pieces of information. First. each macrostate is anchored around a single free energy minimum containing the preponderance of the probability mass of the microstate ensemble. These projections indicate a relatively narrowly peaked distribution of microstates around the free energy minimum with a broad tail. Second, the location of each macrostate free energy minimum in the $d_{3^\prime}^{-1}$-$d_{5^\prime}^{-1}$ projection readily illuminates the base pair pattern within the duplex. Within the fully hybridized H state (blue) $d_{3^\prime}^{-1}$ $\approx$ $d_{5^\prime}^{-1}$ $\approx$ 1.2 nm$^{-1}$. A 2 b.p.\ shift to 3$^\prime$ is manifest in a decrease in $d_{3^\prime}^{-1}$ (increase in $d_{3^\prime}$) and a an increase in $d_{5^\prime}^{-1}$ (increase in $d_{5^\prime}$) producing the stable 3S2 state (green). A further 2 b.p.\ shift to 3$^\prime$ results in a further decrease in $d_{3^\prime}^{-1}$ and also a decrease in $d_{5^\prime}^{-1}$ (back down to the same value as in the H state) to produce the 3S4 state (black). Analogous arguments apply to the 2 b.p.\ 5$^\prime$ shifted state 5S2 (orange) and 4 b.p.\ 5$^\prime$ shifted state 5S4 (purple). The infinitely separated dehybridized state D exists at the origin, but limited by the finite box size corresponds to $d_{3^\prime}^{-1}$ $\approx$ $d_{5^\prime}^{-1}$ $\approx$ 0.2 nm$^{-1}$ (red). The GC-core frayed state F4 (orange) exists as a heart-shaped region between the H and D sates, with each symmetric lobe of the heart corresponding to fraying of one or other of the loose ends. (d) Coloring the microstates by their membership probabilities to their assigned macrostate under a fuzzy clustering procedure shows that microstates lying in the vicinity of the free energy minima have high probabilities whereas those at the macrostate cluster boundaries have less definitive assignments. The former comprise the preponderance of the microstate ensemble while the latter represent intermediate configurations that are transiently occupied as the system transitions from one macrostate to another.}}

@@

\end{figure}

\clearpage
\newpage

\textcolor{blue}{1)      It is unclear how five datasets were selected used for cross-validation in the SRVs (P. 10 L. 43). One would naively think a 50:50 split of a single trajectory would result in only one dataset. A bit more detail could be provided here.}

\textbf{Author Reply:}   Five-fold cross-validation was conducted by constructing five random partitions of the 40 independent trajectories into a training set of 20 trajectories and a validation set of 20 trajectories. This manner of splitting the data at the level of complete trajectories -- as opposed to splitting a single trajectory into multiple subtrajectories -- is desirable as it allows our SRV/MSM approach to learn over long 25 $\mu$s trajectories and infer slowly relaxing collective modes that may not be accessible within short sub-trajectories~\citep{Sidky2019High-ResolutionVAMPnets}. We have better clarified this procedure for the reader by revising the text on p.~11 to read:

@@

We guarded against overfitting using five-fold cross-validation in which we \rood{constructed five random partitions of the 40 independent simulation trajectories into a training set of 20 trajectories and a validation set of 20 trajectories}.\citep{Sidky2019High-ResolutionVAMPnets}

\textcolor{blue}{2)      For the validation of embedding dimensions (P. 11 L. 3), it is hard to judge the presence of a kink at 5D in the AT-all sequence since only a maximum of 6D is shown in Fig.~S1. Could a few additional dimensions be provided?}

textbf{Author Reply:}   We agree that it is more difficult to see the kink for the AT-all sequence than the other sequences. As suggested by the reviewer, we added two additional modes to the cross-validation procedure that does indeed better expose this kink at five slow modes. We have updated the corresponding panel Fig.~S1 and also reproduce it below.

% Fig. S1
\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=1.0\textwidth]{FigS1.pdf}
         \caption*{\textbf{Figure S1.}Five-fold cross-validation of the SRV VAMP-2 scores to select the optimal number of SRV coordinates for each sequence. A knee in the VAMP-2 plot was identified at the fifth, fourth, third, and second slow modes for AT-all, GC-end, GC-core, and GC-mix, respectively. An embedding of corresponding dimensionality was then used to cluster frames into discrete states. The absence of any significant separation in the training and testing VAMP-2 scores demonstrates that model is not overfitted.}
        \label{fig:SIFig1}
	\end{center}
\end{figure}

@@

\textcolor{blue}{4)      In Figure 1 the units for rates are not provided.}

\textbf{Author Reply:}   Thank you for catching this error. Fig.~1 has been updated to include units in s$^{-1}$.

\textcolor{blue}{5)      In the comparison between experiments and theory for slow and fast T-jump IR responses (P. 17 L. 9), were there any correlations between the goodness-of-fit in the calculated rates (Figure S4) and the agreement with experiments for a given system? It is hard to judge as is because the goodness-of-fit values were not provided with Figure S4.}

\textbf{Author Reply:}     We have updated all panels of Fig.~S4 plots to include $R^2$ goodness-of-fit values determined by linear least squares fits of the models in log-space (i.e., $\log \left( f \right) = -k_d t$) and have also re-plotted the data on log-linear axes for easier comparisons between the data and fits. The revised figure is reproduced below. In all cases except the lowest temperature (T$_m$-5) slow response we report good fits with $R^2$ $>$ 0.88. We see marginally less good fits for the GC-core sequence compared to the other three systems, possibly -- as we observe in Section 3.1 of the main text -- due to its high propensity to fray, but the reduction in the quality of the fit is only marginal and in absolute terms the fits are still very good \noteg{not sure what we're reporting here? \rood{X.XX} $<$ $R^2$ $<$ \rood{X.XX})}.

% Fig. S4
\begin{figure}[ht!]
	\centering
    \includegraphics[width=.75\textwidth]{FigS4.pdf}
   \caption*{\textbf{Figure S4.} Exponential fits for both slow (dissociation) and fast (fraying) response for all four sequences during ``computational T-jump'' experiments. From 120 independent 1 $\mu$s simulations, we compiled the slow response data by recording the fraction of sequences with both central Watson-Crick base pairs intact as a function of time, and the fast response data as the fraction of sequences with both terminal Watson-Crick base pairs intact as a function of time. We define a Watson-Crick base pair to be intact if the centers-of-mass of the two comlementary bases lie within a linear distance of 1.3 nm. We extracted our computational estimate of $k_d^\mathrm{fast}$ by fitting a decaying exponential to the fraction of bound A:T termini as a function of time $f_\mathrm{unfrayed}(t) = \exp(-k_d^\mathrm{fast}t)$. Similarly, we extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$. \rood{We report within the figure legend to each panel the coefficient of determination $R^2$ for a least squares linear fit of the model to the data in log space (i.e., $\log \left( f \right) = -k_d t$) and data are plotted on log-linear axes to facilitate visual comparison of the fits. In all cases we observe excellent fits of the models to the data with all $R^2$ values equal to or better than 0.88 (except for the slow response at the the lowest temperature T$_m$-5K where dissociation events are sparse).}}
    \label{fig:SIFig4}
\end{figure}

@@

\clearpage
\newpage

\textcolor{blue}{6)       The authors chose to compute the temperature jump relaxation rates from fits to quantities obtained directly from the MD simulation. However, it is not clear why the authors did not also compute the relaxation of observables from estimated Markov state models following (P. N. A. S. USA. 2011;108(12):4822-4827). This approach has also been previously used to compare kinetics obtained from Markov state models with other spectroscopic temperature-jump based experiments in DNA (J. Phys. Chem. B. 2019;123(10):2291-2304.) and RNA (J. Chem. Theory Comput. 2017;108(2):926-934) oligomers which should be mentioned. If possible, it would be interesting to compare or discuss the differences between their approach and these previous approaches.}

\textbf{Author Reply:}   The reviewer raises an important point regarding the dynamical fingerprinting methodology that have been developed to compare MSM predictions and relaxation experiments. We are familiar with these techniques, are currently applying them in our current work on abasic DNA, and did actually attempt to apply them in the present work. These analyses, however, were less informative than we had hoped due to three primary factors.

@@

``Taken together, our analyses establish new molecular-level understanding of the sequence-dependent kinetics and pathways through quantitative predictive models for the long-time system dynamics, resolution of the dynamical folding pathways and metastable states, and elementary design rules with which to control the dynamical behaviors of the system. We anticipate that this new foundational understanding, and the extension of our approach to more extensive families of DNA sequences, can guide the rational design of DNA oligomers with tailored kinetic properties engineered for DNA nanotechnology applications such as DNA-PAINT. \citep{Shah2019, Strauss2020UpDNA-PAINT}''

\textcolor{blue}{One limitation of the study is its reliance on ``brute-force'' simulations (rather than using advanced sampling/rare-event techniques) to probe the transition. This means that systems of very high concentration (7 M) and temperatures close to the melting point have to be used, which are probably not the conditions of most interest.}

\textbf{Author Reply:}    (I) We agree with the reviewer that it may be possible to use biasing techniques in order to efficiently collect the data required to construct the MSMs. We are very familiar with the use of such techniques: indeed, another major interest of our research group is iterative strategies to perform data-driven determination of collective variables from simulation trajectories, enhanced sampling in these variables, and then estimation of new and improved collective variables from the enhanced sampling trajectories \citep{Sidky2020MolecularSimulators, Sidky2020MachineSimulation, Chen2019CapabilitiesSystems, Chen2019NonlinearVAMPnets, Phys2018CollectiveDesign, Chen2018MolecularExploration}. In the present work, we chose not to employ biased simulations and instead collect unbiased trajectories for three main reasons.

-- First, the 3SPN.2 coarse-grained DNA model is sufficiently inexpensive that it was possible to collect adequate unbiased simulation data to parameterize robust MSMs at reasonable costs. As we state on p.~8:

@@

``Each simulation was conducted for 26 $\mu$s and frames saved to disc every 100 ps. Each simulation required $\sim$24 CPU-hours on 28$\times$Intel E5-2680v4 CPU cores. The first 1 $\mu$s of each run was discarded for equilibration providing us with 40$\times$25 $\mu$s = 1 ms of simulation data for each sequence, during which time we observed 55-100 hybridization/dehybridization events.''

-- Second, the unbiased simulation data are in good agreement with trends in experimental T-jump IR measurements of fast and slow relaxation responses (Section 3.1, Fig.~S\rood{4}) and are sufficient to furnish robust, low-variance, and well-converged MSMs (Section 3.2, Figs.~S\rood{1-3}). This indicates that the data are of sufficient quality and volume to present favorable experimental comparisons and learn reliable long-time kinetic models.

-- Third, although it is straightforward to correct biased simulation data to recover the unbiased thermodynamics, it is more challenging to recover the unbiased kinetics. Although a number of approaches do exist, it has been our experience that the kinetic corrections can present difficulties in numerical convergence. That is not to say that we could not have successfully applied this methods, but just that we did not have cause to do so since the unbiased data were not unreasonably expensive to collect and showed favorable experimental comparisons and produced robust MSMs.

We did note the possibility of performing biased simulations in the Introduction on p.~4-5, and have now elaborated these remarks to better explain our choice to proceed with unbiased simulations:

``The effect of the applied bias upon the thermodynamics can be rigorously corrected for using standard reweighting techniques \citep{Gallicchio2005TemperaturePaths, Souaille2001ExtensionCalculations, Shirts2008StatisticallyStates, Kumar1992THEMethod}. \rood{Rigorous elimination of the bias in the kinetics is critical for the construction of robust and reliable Markov state models that reflect the true system dynamics and are uncontaminated by any residual effects of the biasing potentials used to induce good sampling. A number of approaches to correct the kinetics are also available, including Girsanov reweighting, transition-based reweighting analysis (TRAM), dynamic histogram analysis method (DHAM), and their derivatives~\citep{Prinz2011OptimalDynamics, Chodera2011DynamicalTemperatures, Stelzl2017DynamicSimulations, Donati2017GirsanovModels, Donati2018GirsanovSimulations, Quer2018ANCOORDINATES} \citep{Mey2014XTRAM:States}. The application of these methods under the conditions of high bias necessary for good sampling can, however, present challenges for numerical convergence.} A number of coarse-grained DNA force fields have been developed that enable direct observation of these events over microsecond time scales via unbiased coarse-grained molecular dynamics simulations, \citep{Romano2013DNADependence, Hinckley2013AnHybridization, Maciejczyk2014DNAModel, Markegard2015, Dans2016MultiscaleDNA} which, up to an acceleration factor associated with the smoothing of the underlying free energy landscape inherent to the coarse-graining procedure, can preserve a faithful model of the unbiased dynamics and associated pathways. These models have previously been used to study biological phenomena such as nucleosome dynamics\citep{Lequieu2016Tension-dependentUnwrapping, Lequieu2017InSliding} and transcription factor binding \cite{Terakawa2015P53Simulations, Tan2018DynamicData} as well as nanotechnology applications such as strand displacement \citep{Srinivas2013OnDisplacement, Haley2020DesignDisplacement} and DNA origami \citep{Snodin2019Coarse-grainedOrigami, Doye2020TheOrigami}. \rood{In this work, we choose to employ a coarse-grained model for DNA \citep{Hinckley2013AnHybridization} that is sufficiently inexpensive to enable the collection of sufficient volumes of unbiased simulation trajectories and adequately sample configurational space that we do not need to appeal to biasing strategies to enhance convergence nor apply any \textit{post hoc} corrections to the thermodynamics or kinetics.}''

### 2021-07-27 cover_letter/response_letter_MJ_210721.tex

(II) Our specification of a single strand concentration of 7 mol/L in Section 2.1.1 on p.~7 was an (embarrassing) typo -- the true concentration is actually 7 mmol/L (i.e., 7 mM). This is only 3.5$\times$ higher than the 2 mmol/L experimental concentration, and so is actually quite a good approximation for the experimental conditions. We regret our error, and have corrected this text to now read:

@@

``A single pair of self-complementary sequences were placed in a cubic periodic box with side length 7.8 nm corresponding to a single-strand concentration of \rood{7 mM}. \rood{This concentration is only 3.5$\times$ larger than the 2 mM concentration employed in our experimental analyses (Section 2.2).}''

(III) Finally, we deliberately chose to explore the melting temperature as the state point at which the hybridized and dehybridized populations are in balance and sought to probe the mechanism and kinetics of the hybridization/dehybridization transitions under these equilibrium conditions corresponding to no net thermodynamic preference for either state (i.e., minimally driven conditions). A favorable side effect of this choice was indeed that we observe approximately equal numbers of hybridization and dehybridization events in a single (sufficiently long) simulation, whereas temperatures away from $T_\mathrm{melt}$ would favor one over the other. It is an interesting question to consider how the mechanism and kinetics of hybridization/dehybridization may change with temperature. Exploring small excursions either side of $T_\mathrm{melt}$ may still be possible using unbiased simulations, but we concur with the reviewer that biased sampling would be required to observe sufficient numbers of hybridization (dehybridization) events at large excursions above (below) $T_\mathrm{melt}$.

In response to this point, we have modified text in the terminal paragraph of the Conclusions on p.~38 observing that this may be an interesting point for follow-up study:

``Going forward, we will extend this work to discern more general trends in sequence-dependent hybridization/dehybridization for a wider range of oligomer sequences and motivate strategies for experimental comparisons. \rood{We also suggest that the MSM approach followed in this work, possible coupled with biased sampling and reweighting techniques \citep{Wu2014StatisticallyStates, Scherer2015PyEMMAModels} , may be well-suited to exposing changes in the hybridization/dehybridization mechanism and kinetics as a function of temperature.}''

@@

\textcolor{blue}{Very small points:
+ p16 line 3: "Although the 3SPN.2 model reproduces melting temperatures relatively well, we observed a systematic 4 K under-prediction relative to experiment and so we apply a universal (+4) K corrective calibration to our computational results". This sentence confused me at first. I presume you are just shifting your simulated kd slow curves to better match experiment, rather than shifting the data so that the melting temperatures agree. You may want to be more explicit as I read it as the latter at first. (Note one shouldn't expect the melting temperatures to agree as very different concentrations are being used in simulation and experiment, and one also has to correct for finite-size effects when estimating melting points from single duplex simulations (https://doi.org/10.1088/0953-8984/22/10/104102)}

\textbf{Author Reply:}   (I) We regret our lack of clarity in our choice of wording. Yes, we simply shifted the $k_d$ curves by a universal temperature shift, determined to be +4 K, that produced the optimal match with the experimental data.

We have also reworded the unclear wording regarding the temperature correction in Section 3.1 p.~16 to now read:

``Although the 3SPN.2 model reproduces melting temperatures relatively well, we observed a systematic 4 K under-prediction relative to experiment and so we apply a \rood{universal (+4) K corrective temperature shift} to our computational results.''

@@

(II) We regret our erroneous reporting of a 7 M rather than 7 mM concentration as detailed in our previous response. As we have now clarified, our simulations and experiments are conducted at similar concentrations (7mM and 2 mM, respectively). We agree that we still cannot expect perfect agreement in the melting temperatures due to the residual small  difference in concentrations, but likely this is small in comparison to discrepancies introduced by the approximations inherent in the coarse-grained molecular model.

(III) We greatly appreciate this comment and suggestion to apply a fluctuation correction to our data in order to correct for artifacts associated with finite-size effects following the approach detailed in Ref.~\citep{Ouldridge2010ExtractingSimulations}. We have reanalyzed our slow repsonse data under the homodimer correction and this has pleasingly resulted in a slightly improved agreement between our simulations and experiment. The revised figure is reproduced below. \noteg{We note that it is unclear how this correction should be applied to a periodic system that is not in thermodynamics equilibrium, however we choose to apply the correction so that we will observe asymptotically correct behavior at long timescales.}

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Fig1.pdf}
        \caption*{\textbf{Figure 1.} Experimental measurements and computational predictions of slow and fast at T-jump IR responses. Results are reported in terms of the final T-jump temperature. (a) The experimental and simulated slow rate constants $k_d^\mathrm{slow}$ corresponding to duplex dissociation over long time scales. \rood{We apply a fluctuation correction proposed by Ouldridge et al.\citep{Ouldridge2010ExtractingSimulations} to account for finite-size effects and ensure an asymptotically correct dimer fraction} (b) The experimental and simulated fast rate constant $k_d^\mathrm{fast}$ corresponding to terminal base-pair fraying on short time scales. The simulation results are corrected by a sequence-independent scaling factor that corrects for a 10$\times$ acceleration of the slow dissociation dynamics and 120$\times$ acceleration of the fast fraying dynamics. The simulated temperature in all cases is subjected to a (+4) K corrective calibration to account for an observed systematic under-prediction of the melting temperature by the 3SPN.2 model}
        \label{fig:relaxation-comparison}
	\end{center}
\end{figure}

In light of this modification, we have also added the following text to the Section 3.1 on p.~16:

\rood{To correct for finite-size effects associated with the simulation of a single duplex within a periodic simulation box, we apply the fluctuation correction proposed by Ouldridge et al.\citep{Ouldridge2010ExtractingSimulations} to correct our predictions for $f_\mathrm{hybridized}$ bulk yields. We correct the values observed in our simulations $f_\mathrm{hybridized}^\mathrm{sim}$ to the bulk corrected result $f_\mathrm{hybridized}$ using the following relations:}

    \begin{equation}\label{}
    \phi = \frac{\mathrm{f_\mathrm{hybridized}^\mathrm{sim}}}{\mathrm{1 - f_\mathrm{hybridized}^\mathrm{sim}}}
    \end{equation}

    \begin{equation}\label{}
	f_\mathrm{hybridized} = (1+\frac{1}{4\phi}) - \sqrt{(1+\frac{1}{4\phi})^2 - 1}
	\end{equation}

Finally, we have also edited the following text in Section 3.1 on p.~17 to clarify that the fluctuation correction is only applied for the purposes of the experimental comparison in Section 3.1 and is not propagated throughout the remainder of the paper:

@@

\textbf{Author Reply:}   We thank the reviewer for this suggestion and regret our lack of clarity in this matter. All simulations performed in this work were conducted at the melting temperature, with the exception of the simulations discussed in Section 3.1 where we consider a range of temperatures in order to draw comparisons to experimental measurements over this temperature range. We have added a number of clarifications to the text in order to more clearly convey this important point to the reader.

In Section 2.1.1 of the Methods on p.~7, we have added:

``\rood{With the exception of the simulation data reported in Section 3.1 where we draw comparisons against experimental data collected over a range of temperatures,} each sequence was simulated at its melting temperature -- AT-all: 309 K, GC-end: 317 K, GC- core: 324 K, GC-mix: 324 K -- in order to maximize the number of spontaneous transitions between dissociated and hybridized states.''

In Section 2.1.2 of the Methods on p.~8, we have added:

``In this work, MSMs were constructed for each of the four DNA sequences \rood{at their respective melting temperatures} from the 40$\times$25 simulation trajectories following a four-step protocol...''

In the first sentence of Section 3.3 on p.~21, we have added:

@@

``Thermodynamic and kinetic predictions of the sequence-specific MSMs \rood{fitted at the sequence melting temperatures} for AT-all, GC-end, GC-core, and GC-mix.''

In the first sentence of the caption of Fig.~2, we have added:

``Comparison of the macrostate free energy predictions of the MSMs and nearest neighbor (NN) thermodynamic model \rood{at the sequence melting temperatures}.''

In the first sentence of Section 3.4 on p.~26, we have added:

@@

``In addition to thermodynamic stabilities, the macrostate MSM also furnishes quantitative and interpretable predictions of hybridization and dehybridization pathways and mechanisms \rood{at the sequence melting temperatures}.''

In the second sentence of the Conclusions in Section 4 on p.~36, we have added:

@@

``We conducted 1 ms of unbiased coarse-grained molecular dynamics simulations \rood{at the melting temperature of each sequence} and...''

\textcolor{blue}{2. P7 line 13 - surely these simulations were not really done at 7M? That sounds impossibly dense.}

\textbf{Author Reply:}   We regret our (embarrassing) typo -- the concentration is in fact 7 mM, not 7 M. The calculation is below and we have made the correction to Section 2.1.1 on p.~7. We regret our error.

@@

\begin{equation}\label{}
\frac{2 \mathrm{\:molecules}}{\mathrm{box\:volume}} = \frac{2 \mathrm{\:molecules}}{(7.80 \times 7.80 \times 7.80) \mathrm{\:nm}^3} = \frac{3.32 \times 10^{-24} \mathrm{\:moles}}{4.75 \times 10^{-25} \mathrm{\:m}^3} = 7.0 \frac{\mathrm{moles}}{\mathrm{m}^3} = 7.0 \frac{\mathrm{millimoles}}{\mathrm{L}} = 7 \; \mathrm{mM}
\end{equation}

\textcolor{blue}{3.  The authors repeatedly refer to "slithering" mechanisms for rearrangement. I believe this is unfortunate. In the context of the 3SPN model, the "slithering" term was first introduced in Ref. 49 by Sambriski et al, and describes two helical single strands sliding along each other. As noted by Hinckley et al in Ref 18, that mechanism is largely an artifact of the form of the potential in 3SPN.1. The rearrangement mechanism in 3SPN.2 is more of a "defect diffusion", analogous to the "inchworm mechanism" previously identified in the oxDNA model (ref 52 - note, the author list for this reference is wrong). Given this context, it seems unnecessarily confusing to talk about "slithering".}

\textbf{Author Reply:}    (I) We share the reviewer's concerns with using ``slithering'' to describe a process that is more appropriately referred to as ``defect diffusion'', ``inchworm mechanism'', or ``out-of-register diffusion''.

In Section 2.1.2 on p.~9, we have modified the text to replace ``slithering'' with ``out-of-register diffusion'':

@@

``An energy disconnectivity graph-based approach was used to interrogate the differences in hybridization rates and mechanisms between GGGGGG and GCGCGC hexamers to reveal strong deviations from ``all-or-nothing'' behaviors and the importance of zippering and \rood{out-of-register diffusion mechanisms.~\citep{Xiao2019}''}

The only other place we use the term is in Section 3.4 on p.~29 where we refer to work conducted in Ref.~\citep{Xiao2019}. Here we choose to maintain the term ``slithering'' to maintain consistency with the mechanism nomenclature adopted in this paper, but place it in quotation marks to indicate that we are quoting from that paper and it is not a term that we would necessarily favor in the context of our own work.

@@

(II) We also appreciate the reviewer pointing out the incorrect author list in ref 52, and we have fixed the issue.

\textcolor{blue}{4. P15: I am a bit surprised that a separation of 1.3nm between the central base pairs is enough to guarantee strand separation, given that frayed end can separate by 1.3nm when the core is intact. Can the authors verify that all their duplexes had truly separated at this point?}

\textbf{Author Reply:}   This a good point of clarification. We motivated our choice of this distance as that beyond which the strands are effectively non-interacting, but also verified that our estimates of $k_d^{slow}$ are robust to choices of this cutoff in the range 0.8--2.5 nm. We illustrate this in \textbf{Figure R1} below where we show that our extracted values of $k_d^{slow}$ are robust in changing the cutoff from 1.3 nm to 2.0 nm.

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{FigR1.pdf}
        \caption*{\textbf{Figure R1.} Comparing the experimental slow response when increasing the cutoff parameter from 1.3 nm to 2.0 nm}
        \label{fig:compare_cutoffs}
	\end{center}
\end{figure}

We have modified the relevant text in Section 3.1 on p.~15 to read:

@@

``First, we tracked the slow response corresponding to duplex dissociation in our simulations by compiling the distribution of times at which both of the central base pairs first separate to a distance of 1.3 nm starting from an initial fully hybridized duplex. This cutoff was selected as the distance beyond with the strands are effectively non-interacting and define the dissociated state. \rood{We verified that our results were robust to choices of this cutoff over the range 0.8-2.5 nm.}''

\textcolor{blue}{5. The authors refer to "exponential" increase in Fig. 1. However, it's hard to see that an increase is exponential by eye unless the data is plotted on a log scale.}

\textbf{Author Reply:}   We agree that a log-linear plot would better expose these trends. We have replotted Fig.~1 and Fig.~S\rood{4} on log-linear axes and reproduce these below.

\begin{figure}[ht!]
	\begin{center}
        \includegraphics[width=\textwidth]{Fig1.pdf}
        \caption*{\textbf{Figure 1.} Experimental measurements and computational predictions of slow and fast at T-jump IR responses. Results are reported in terms of the final T-jump temperature. (a) The experimental and simulated slow rate constants $k_d^\mathrm{slow}$ corresponding to duplex dissociation over long time scales. \rood{We apply a fluctuation correction proposed by Ouldridge et al.\citep{Ouldridge2010ExtractingSimulations} to account for finite-size effects and ensure an asymptotically correct dimer fraction} (b) The experimental and simulated fast rate constant $k_d^\mathrm{fast}$ corresponding to terminal base-pair fraying on short time scales. The simulation results are corrected by a sequence-independent scaling factor that corrects for a 10$\times$ acceleration of the slow dissociation dynamics and 120$\times$ acceleration of the fast fraying dynamics. The simulated temperature in all cases is subjected to a (+4) K corrective calibration to account for an observed systematic under-prediction of the melting temperature by the 3SPN.2 model}
        \label{fig:relaxation-comparison}
	\end{center}
\end{figure}

% Fig. S4
\begin{figure}[ht!]
	\centering
    \includegraphics[width=.75\textwidth]{FigS4.pdf}
   \caption*{\textbf{Figure S4.} Exponential fits for both slow (dissociation) and fast (fraying) response for all four sequences during ``computational T-jump'' experiments. From 120 independent 1 $\mu$s simulations, we compiled the slow response data by recording the fraction of sequences with both central Watson-Crick base pairs intact as a function of time, and the fast response data as the fraction of sequences with both terminal Watson-Crick base pairs intact as a function of time. We define a Watson-Crick base pair to be intact if the centers-of-mass of the two comlementary bases lie within a linear distance of 1.3 nm. We extracted our computational estimate of $k_d^\mathrm{fast}$ by fitting a decaying exponential to the fraction of bound A:T termini as a function of time $f_\mathrm{unfrayed}(t) = \exp(-k_d^\mathrm{fast}t)$. Similarly, we extracted our computational estimate of $k_d^\mathrm{slow}$ by fitting a decaying exponential to the fraction of hybridized sequences as a function of time $f_\mathrm{hybridized}(t) = \exp(-k_d^\mathrm{slow}t)$. \rood{We report within the figure legend to each panel the coefficient of determination $R^2$ for a least squares linear fit of the model to the data in log space (i.e., $\log \left( f \right) = -k_d t$) and data are plotted on log-linear axes to facilitate visual comparison of the fits. In all cases we observe excellent fits of the models to the data with all $R^2$ values equal to or better than 0.88 (except for the slow response at the the lowest temperature T$_m$-5K where dissociation events are sparse).}}
    \label{fig:SIFig4}
\end{figure}

@@

\clearpage
\newpage

\textcolor{blue}{6. Do the authors have any data on the scaling of k\_hybridization with T? In experiment, rates appear to increase with temperature (albeit with a much weaker dependence than de-hybridization rates). Simple "nucleation and zipper" models would predict the opposite, as does the oxDNA model (see ref 52).}

\textbf{Author Reply:}   This is an interesting question and are, in fact, considering embarking upon some additional work in this vein. We do not, however, have any data on this scaling at the current time. Rigorous engagement of this question would require us to fit MSMs at a variety of temperatures in order to compare the pathways, mechanisms, and rates for hybridization and dehybridization. In the present work, we have only constructed MSMs at the melting temperature of each sequence and the collection of sufficient data to parameterize MSMs at other temperatures would require both (i) a substantial investment of time to collect the $\sim$1 ms of data required to fit robust MSMs and (ii) the use of biasing techniques in order to adequately sample hybridization (dehybridization) events at large temperature excursions above (below) the melting temperature. As such, we do not have the necessary data to engage this question in the present work, but would like to explore it in the future.

Reviewer \#2 raised a related point in their Comment \#2 and we have made the following modification to the terminal paragraph of the Conclusions on p.~38 observing that this may be an interesting follow-up study:

``Going forward, we will extend this work to discern more general trends in sequence-dependent hybridization/dehybridization for a wider range of oligomer sequences and motivate strategies for experimental comparisons. \rood{We also suggest that the MSM approach followed in this work, possibly coupled with biased sampling and reweighting techniques \citep{Wu2014StatisticallyStates, Scherer2015PyEMMAModels}, may be well-suited to exposing changes in the hybridization/dehybridization mechanism and kinetics as a function of temperature.}''

@@

\textcolor{blue}{7. The authors could explain a bit more carefully (perhaps in the SI) how they handle the multiple nearest-neighbour states that correspond to their microstates (eg. F4 clearly involves more than one NN configuration).}

We have modified the text in the ``Nearest neighbor model of duplex thermodynamics'' text in the Supporting Information to better describe for the reader how we performed this calculation. Specifically, we have appended the following text at the end of this discussion:

\noteg{Have not incorporated this yet, but this would be one way to address it without making the FE calculations significantly (and unnecessarily) more complicated}
``\rood{In applying the NN models to each macrostate, we made the simplifying assumption that the ensemble of microstates constituting each macrostate could be represented by a single pattern of Watson-Crick base pairing that are schematically illustrated in Fig.~2a. Indeed, each macrostate is composed of an ensemble of partially frayed stayed with higher free energy contributions than the intact pattern. However, when looking at the difference between states, we have found that incorporating frayed contributions changes 
$\Delta$ F values by $<$ 1 kJ/mol when compared to used a single macrostate. Given the extent to which a single NN representation captures each macrostates, we have elected to maintain a single microstate free energy estimate for each macrostate}''
